---
title: "Analysis of larval and spat responses"
author: "Ariana S Huffmyer and Shayle B Matsuda"
date: "02/14/2022"
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---
Data analysis for Alexander et al., "Larval thermal conditioning does not improve post-settlement thermal tolerance in the dominant reef-building coral, *Montipora capitata*".    

# Set Up  

Load required libraries.  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library("car") #levenes test
library("plotrix") #plotting
library("reshape2") #data shaping
require("gridExtra") #Arrange Plots for output
require("utils") #R utilities
library("tidyverse") #data tidy
library("ggplot2") #plotting
library("lsmeans") #post-hoc tests
library("effects") #plot effects of modeling
library("lme4") #linear mixed modeling
library("lmerTest") #calculate p-values in models
library("emmeans") #post-hoc tests
library("cowplot") #arrange plots
library("multcomp") #arrange plots
library("blmeco") #dispersion tests 
```

# Larval survivorship  

Load datasets for larval survivorship.   
```{r}
#load and manipulate data sets
larvae<-read.csv("Data/surv_larvae.csv", header=T, na.strings="NA") #load larval survivorship dataset
larvae$ConicalVol<-as.numeric(larvae$ConicalVol) 

larvae$perML<-larvae$Larvae/larvae$Volume #calculate larval concentration as larvae per mL seawater
larvae$total<-larvae$perML*larvae$ConicalVol #calculate total number of larvae by multiplying conical volume by larval concentration
```

Calculate percent change in larval concentration relative to starting concentration at each subsequent time point: (larvae end - larvae start) / larvae start  

Due to this method, it is possible for survival values to be >100%, due to variation in larval concentration samples.  

```{r}
#first, summarize by conical to make comparisons  
larvae_sum<- plyr::ddply(larvae, c("Conical", "Day", "Treatment"), summarise, 
                           mean = mean(perML, na.rm=TRUE)) #calculate mean larval concentration

larvae_sum<-larvae_sum%>%
  filter(., !Day=="2") #remove a day with incomplete measurements

#Calculate proportion survival
z <- larvae_sum %>%
  group_by(Conical) %>% #group by conical
  arrange(Day, .by_group = TRUE) %>% #arrange by date within conical
  mutate(pct_change = ((mean-first(mean))/first(mean)), prop_change = (mean/first(mean))) #calculate percent change and proportion change relative to the first timepoint
```

Generate summary table of larval survivorship.  
```{r}
#generate a summary table of larval density over time (days 0, 1, 3, 4) in ambient, cool, and high treatments
larvae_surv_table <- plyr::ddply(z, c("Treatment", "Day"), summarise, 
                 N    = length(prop_change[!is.na(prop_change)]), 
                 mean = mean(prop_change, na.rm=TRUE) * 100, 
                 sd   = sd(prop_change, na.rm=TRUE) * 100, 
                 se   = sd / sqrt(N), 
                 max = max(prop_change, na.rm=TRUE), 
                 lower = mean-se, 
                 upper = mean+se 
)
larvae_surv_table 
```

Analyze larval density over time between treatments with a linear mixed model with conical nested within treatment as a random effect due to repeated measures.  

Build model: `lmer(prop_change~Treatment + Day  + Treatment:Day + (1|Treatment:Conical), data=z)`
```{r}
z$Day<-as.factor(as.character(z$Day))
model1<-lmer(prop_change~Treatment + Day  + Treatment:Day + (1|Treatment:Conical), data=z) #model
```

View summary and significance of main effects through a Type II ANOVA analysis.  
```{r}
summary(model1) #summary of model
anova(model1, type="II") #anova of model
```

Day and Treatment:Day are significant.  

Check residual assumptions for the model.  
```{r}
hist(residuals(model1)) 
qqPlot(residuals(model1)) 
leveneTest(residuals(model1)~Treatment * as.factor(Day), data=z) #passes
```

Conduct Tukey posthoc test of model.  
```{r}
posthoc1<-emmeans(model1, ~Treatment|Day, adjust="Tukey") #examine posthoc comparisons for each day across treatments
cld(posthoc1)
pairs(posthoc1)
```

Plot larval survival.  
```{r}
#graph larval density over time
LarvalSurvPlot<-ggplot(data=larvae_surv_table, aes(x=Day, y=mean, colour=Treatment)) + 
  scale_colour_manual(name="Larval Treatment",
                    values=c("lightblue", "darkblue", "red"),
                    labels=c("Ambient", "Cool", "Warm"))+
  geom_line(position=position_dodge(0.2), size=1) + 
  geom_point(size=3, position=position_dodge(0.2)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.0, size=1, position=position_dodge(0.2), linetype=1)+ 
  theme_classic()+ 
  ylim(0,135)+
  theme(text = element_text(size = 18, color="black"))+ 
  theme(axis.text = element_text(size = 18, color="black"))+ 
  theme(legend.title = element_text(size = 18, color="black", face="bold"))+
  theme(legend.text = element_text(size = 18, color="black"))+
  theme(axis.title = element_text(size = 18, color="black"))+ 
  theme(legend.position = "none")+ 
  theme(plot.margin = margin(1, 0.1, 0, 0.1, "cm")) +
  ylab(expression(bold(paste("Larval Survival (%)")))) + 
  geom_text(x=1.5, y=10, label="p(Larval Treatment x Day)=0.009", size=6, color="black") +
  geom_text(x=3.95, y=0, label="*", size=10, color="darkblue") + #day4 cool
  xlab(expression(bold("Days")));LarvalSurvPlot 
```

# Larval settlement  

Load data for larval settlement. Calculate proportion of larvae settled in each chamber. Each chamber was loaded with a known number of larvae at the start of the settlement period. At each time point the number of spat were counted and proportion settlement is calculated as the number of settled spat at each time point divided by the number of starting larvae.     
```{r}
#load dataset
settle<-read.csv("Data/settlement.csv", header=T, na.strings="NA") #load data 

#calculate total settlement by chamber by adding the number of settlers across all three plugs in each chamber
total_settle <- plyr::ddply(settle, c("Day", "Chamber"), summarise, 
                 total    = sum(Total, na.rm=TRUE), 
                 aggregate = sum(Aggregate, na.rm=TRUE), 
                 individual   = sum(Individual, na.rm=TRUE) 
)

#merge in identifying information of tank, treatment, and number of starting larvae
total_settle$Tank<-settle$Tank[match(total_settle$Chamber, settle$Chamber)]
total_settle$Treatment<-settle$Treatment[match(total_settle$Chamber, settle$Chamber)]
total_settle$Starting.Larvae<-settle$Starting.Larvae[match(total_settle$Chamber, settle$Chamber)]

#calculate proportion settled in each chamber
total_settle$propSettled<-total_settle$total/total_settle$Starting.Larvae
total_settle$propAggregate<-total_settle$aggregate/total_settle$total
```

Generate a summary table of larval settlement over time originating from each larval treatment - settlement is shown as number of larvae settled per 100 larvae.  

```{r}
larval_settle_table <- plyr::ddply(total_settle, c("Treatment", "Day"), summarise, 
                 N    = length(total[!is.na(total)]), 
                 mean = mean(total, na.rm=TRUE), 
                 sd   = sd(total, na.rm=TRUE), 
                 se   = sd / sqrt(N), 
                 max = max(total, na.rm=TRUE), 
                 lower = mean-se, 
                 upper = mean+se 
)
larval_settle_table 
```

Analyze total settlement over time between treatments with a linear mixed model. Because this is count data we will use a poisson distribution in a glmer model. Nest chamber within tank as this is repeated measures.  

Build model: `glmer(total ~ Treatment + Day  + Treatment:Day + (1|Tank/Chamber), data=total_settle, family=poisson)` 
```{r}
total_settle$Day<-as.factor(as.character(total_settle$Day))
model2<-glmer(total ~ Treatment + Day  + Treatment:Day + (1|Tank/Chamber), data=total_settle, family=poisson) 
```

View model summary and significance of main effects using a Type II ANOVA analysis.  
```{r}
summary(model2) 
Anova(model2, type=2)
```

Check normality assumption for model.  
```{r}
qqPlot(residuals(model2)) 
hist(residuals(model2)) 
```

Conduct Tukey posthoc analyses to look for pairwise differences in settlement between treatments at each day.  
```{r}
posthoc2<-emmeans(model2, ~Treatment|Day, adjust="Tukey")
cld(posthoc2)
pairs(posthoc2)
```

Plot larval settlement rates over the settlement period for each temperature treatment.   
```{r}
#graph larval settlement over time (days 1, 2, 3, 5, 6, 7) originating from each larval treatment (ambient, cool, high)
LarvalSettlePlot<-ggplot(data=larval_settle_table, aes(x=Day, y=mean, colour=Treatment)) + 
  scale_colour_manual(name="Larval Treatment",
                       values=c("lightblue", "darkblue", "red"), 
                       labels=c("Ambient", "Cool", "Warm"))+ 
  geom_line(position=position_dodge(0.2), size=1) + 
  geom_point(size=3, position=position_dodge(0.2)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.0, size=1, position=position_dodge(0.2), linetype=1)+ 
  theme_classic()+ 
  ylim(-5,100)+
  theme(text = element_text(size = 18, color="black"))+ 
  theme(axis.text = element_text(size = 18, color="black"))+ 
  theme(legend.text = element_text(size = 18, color="black"))+ 
  theme(legend.position = "none")+ 
  theme(legend.title = element_text(size = 18, color="black", face="bold"))+ 
  theme(axis.title = element_text(size = 18, color="black"))+ 
  theme(plot.margin = margin(1, 0.1, 0, 0.1, "cm")) +
  ylab(expression(bold(paste("Larvae Settled (%)")))) + 
  geom_text(x=4, y=75, label="p(Larval Treatment x Day)<0.001", size=6, color="black") + #obtained from analysis below  
  geom_text(x=1.05, y=18, label="**", size=10, color="red") + #day1 high
  geom_text(x=1.05, y=-6, label="*", size=10, color="darkblue") + #day1 cool
  geom_text(x=2.05, y=19, label="**", size=10, color="red") + #day2 high
  geom_text(x=2.05, y=-4, label="*", size=10, color="darkblue") + #day2 high
  geom_text(x=3.05, y=26, label="**", size=10, color="red") + #day3 high
  geom_text(x=3.05, y=-4, label="**", size=10, color="darkblue") + #day3 cool
  geom_text(x=6.05, y=30, label="**", size=10, color="red") + #day6 high
  xlab(expression(bold("Days"))); LarvalSettlePlot 
```

# Spat survival  
 
Load data for spat survival. Calculate percent spat survival as the number of spat alive at each time point divided by the number of spat alive at the start of the experiment. 

In this analysis, the number of spat alive at each time point is recorded as the number of "successes". The number of larvae that died are "failures". Recruits are tracked by "plug ID". Note that spat are referred to as "recruits" in the code, but spat is used in revised manuscript.   
```{r}
#load and manipulate data
recruits<-read.csv("Data/surv_recruits.csv", header=T, na.strings="NA") #load recruit survivorship dataset
recruits$Tank<-as.factor(recruits$Tank) 

#calculate proportion survival 
recruits$proportion<-(recruits$Success/(recruits$Success+recruits$Failure))*100 #calculate number alive (successes) at each time point divided by the total number of starting recruits (successes + failures).  

#add in larval temperature treatment information - plug 98 did not have larval treatment information and was removed
recruits$Larval.Treatment<-settle$Treatment[match(recruits$Plug.ID, settle$Plug)]
```

Generate a summary table of spat survival over time (days 0, 1, 3, 4) in ambient and high conditions originating from ambient, cool, and high larval treatments treatments. In this experiment recruits that originated from each larval treatment were exposed to ambient and high recruit treatments. Note that "juvenile" terms here refer to spat stage. Terminology revised for manuscript text.      
```{r}

recruit_surv_table <- plyr::ddply(recruits, c("Days", "Juv.Treatment", "Larval.Treatment"), summarise, 
                           N    = length(proportion[!is.na(proportion)]), 
                           mean = mean(proportion, na.rm=TRUE), 
                           sd   = sd(proportion, na.rm=TRUE), 
                           se   = sd / sqrt(N), 
                           max = max(proportion, na.rm=TRUE), 
                           lower = mean-se, 
                           upper = mean+se 
);recruit_surv_table 

recruit_surv_table2 <- plyr::ddply(recruits, c("Days", "Juv.Treatment", "Tank"), summarise, 
                                  N    = length(proportion[!is.na(proportion)]), 
                                  mean = mean(proportion, na.rm=TRUE), 
                                  sd   = sd(proportion, na.rm=TRUE), 
                                  se   = sd / sqrt(N), 
                                  max = max(proportion, na.rm=TRUE), 
                                  lower = mean-se, 
                                  upper = mean+se 
);recruit_surv_table2 

```

Analyze spat survival over time between treatments with a binomial mixed model with a success and failure vector. Plug included as a random effect due to repeated measures. 

Build model for spat survival: `glmer(y~Juv.Treatment * Larval.Treatment * Days + (1|Tank/Plug.ID), family=binomial, data=recruits)`
```{r}
recruits$Days<-as.factor(as.character(recruits$Days))
attach(recruits)

y<-cbind(Success, Failure) #generate binomial vector

model3<-glmer(y~Juv.Treatment * Larval.Treatment * Days + (1|Tank/Plug.ID), family=binomial, data=recruits) 

detach(recruits)
```

View summary of model and significance of main effects with a Type II ANOVA analysis.  
```{r}
summary(model3) 
Anova(model3, type=2)
```

Run Tukey posthoc analysis to look at differences between spat treatments for each day because day*spat treatment was significant, with no effect of larval treatment.  
```{r}
posthoc3<-emmeans(model3, ~Juv.Treatment|Days, adjust="Tukey")
cld(posthoc3)
pairs(posthoc3)
```

Check dispersion assumption.  
```{r}
dispersion_glmer(model3) #no evidence of overdispersion as value is <1.4
```

We now need to account for the effect of colony type (individual vs aggregate) on survival. We found variation in settlement behavior with spat settling both as individuals and aggregates of multiple larvae. The number of aggregates and individuals was recorded and the survival of each individual tracked.    

The success and failure columns reflect the total survival (aggregates + individuals). Generate a new success and failure column for individuals and aggregates separately. Display mean proportion aggregation for each treatment combination.  
```{r}
#success is the number of alive colonies (Ind or Agg columns)
#failure is the number of starting alive colonies - number of alive colonies at a given time point
agg_ind<-recruits%>%
  mutate(Success_Ind=Individual)%>%
  mutate(Success_Agg=Aggregate)%>%
  group_by(Plug.ID)%>%
  arrange(Plug.ID)%>%
  mutate(Failure_Ind=lag(first(Success_Ind), default = first(Success_Ind))-Success_Ind)%>%
  mutate(Failure_Agg=lag(first(Success_Agg), default = first(Success_Agg))-Success_Agg)

#calculate proportion of colonies that are aggregates for these plugs
agg_ind<-agg_ind%>%
  mutate(prop_agg=Aggregate/Total)

agg_summary<-agg_ind%>%
  group_by(Juv.Treatment, Larval.Treatment)%>%
  summarise(mean=mean(prop_agg, na.rm=TRUE));agg_summary

plug_summary<-agg_ind%>%
  group_by(Plug.ID)%>%
  summarise(mean=mean(prop_agg, na.rm=TRUE))%>%
  filter(!is.na(mean))
```

View distribution of proportion aggregation.  
```{r}
hist(plug_summary$mean)
```

There are clear outliers in the histogram for proportion aggregation. Look at distrubution of plugs by proportion aggregation.  
```{r}
n=sum(with(plug_summary, mean<0.5))
d=sum(with(plug_summary, mean<1))
n/d

n=sum(with(plug_summary, mean<0.2))
d=sum(with(plug_summary, mean<1))
n/d

n=sum(with(plug_summary, mean<0.1))
d=sum(with(plug_summary, mean<1))
n/d
```

97% of plugs have <50% aggregates  
80% of plugs have <20% aggregates   
67% of plugs have <10% aggregates  

A majority of plugs have <10% aggregates. Because 97% of plugs have <50% aggregates, we will remove any plugs with >50% aggregation.  

```{r}
agg_ind<-agg_ind%>%
  filter(prop_agg<0.5)
```

Test for differences in aggregation by treatments with an ANOVA.    
```{r}
summary(aov(prop_agg~Juv.Treatment*Larval.Treatment, data=agg_ind))
```

There is a significant effect of treatment on aggregation, therefore this needs to be accounted for in our analysis. Because our experimental design was not designed to examine aggregate vs individual dynamics, we will account for aggregation as a random effect in our modeling. Model structure will be the same as previous model, but with the added random effect of proportion aggregation for each replicate (plug).  

Model: `glmer(y~Juv.Treatment * Larval.Treatment * Days + (1|Tank/Plug.ID) + (1|prop_agg), family=binomial, data=agg_ind)`
```{r}
agg_ind$Days<-as.factor(as.character(agg_ind$Days))
attach(agg_ind)

y<-cbind(Success, Failure)

model4<-glmer(y~Juv.Treatment * Larval.Treatment * Days + (1|Tank/Plug.ID) + (1|prop_agg), family=binomial, data=agg_ind) 

detach(agg_ind)
```

View summary of model and significance of main effects with a Type II ANOVA analysis.  
```{r}
summary(model4) 
Anova(model4, type=2)
```
Results show the same significant effects as the previous model - spat treatment and day interaction is significant. This analysis accounts for variation in settlement behavior.   

Examine Tukey posthoc differences between spat treatments at each day.  
```{r}
posthoc4<-emmeans(model4, ~Juv.Treatment|Days, adjust="Tukey")
cld(posthoc4)
pairs(posthoc4)
```

Examine dispersion assumption.  
```{r}
dispersion_glmer(model4) #no evidence of overdispersion as value is <1.4
```

Graph proportion survival over time in spat.     
```{r}
RecruitSurvPlot<-ggplot(data=recruit_surv_table, aes(x=Days, y=mean, shape=Juv.Treatment, linetype=Juv.Treatment, colour=Larval.Treatment, group=interaction(Juv.Treatment, Larval.Treatment))) + 
  scale_colour_manual(name="Larval Treatment",
                      values=c("lightblue", "darkblue", "red"),
                      labels=c("Ambient", "Cool", "Warm"))+
  geom_line(position=position_dodge(0.3), size=1) + 
  geom_point(size=5, position=position_dodge(0.5), show.legend=FALSE) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.0, size=1, position=position_dodge(0.5), linetype=1)+ 
  scale_linetype_manual(name="Spat Treatment", values=c("solid", "dotted"), labels=c("Ambient", "High"))+
  scale_shape_manual(name="Spat Treatment", values=c(19,21))+ 
  theme_classic()+ 
  ylim(0,100)+
  xlim(0,50)+
  theme(text = element_text(size = 18, color="black"))+ 
  theme(axis.text = element_text(size = 18, color="black"))+ 
  theme(legend.title = element_text(size = 18, color="black", face="bold"))+
  theme(legend.text = element_text(size = 18, color="black"))+
  theme(axis.title = element_text(size = 18, color="black"))+ 
  theme(legend.position = "bottom")+ 
  ylab(expression(bold(paste("Spat Survivorship (%)")))) + 
  theme(plot.margin = margin(1, 0.1, 0, 0.1, "cm")) +
  geom_text(x=28, y=100, label="**", size=10, color="black") +
  geom_text(x=47, y=93, label="**", size=10, color="black") +
  geom_text(x=25, y=25, label="p(Spat Treatment x Day)<0.001", size=6, color="black") +
  xlab(expression(bold("Days Exposure")));RecruitSurvPlot 
```

# Generate final figure  
  
```{r}
#generate figure with larval survival, settlement, and recruit survival in one panel 

figure_surv<-plot_grid(LarvalSurvPlot, LarvalSettlePlot, labels = c("a", "b"), ncol=2, nrow=1, rel_heights= c(1,1), rel_widths = c(1,1), label_size = 20, label_y=1, align="h")

figure_surv_b<-plot_grid(figure_surv, RecruitSurvPlot, labels = c("", "c"), ncol=1, nrow=2, rel_heights= c(1,1), rel_widths = c(1,1), label_size = 20, label_y=1)

ggsave(filename="Figures/DataFigure.pdf", plot=figure_surv_b, dpi=500, width=12, height=10, units="in")
ggsave(filename="Figures/DataFigure.png", plot=figure_surv_b, dpi=500, width=12, height=10, units="in")

```

