---
title: "Temperature analysis"
author: "Ariana S Huffmyer"
date: "6/24/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Set Up  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library("car") #levenes test
library("ggplot")
library("ggplot2") #plotting
library("plotrix") #plotting
library("reshape2") #data shaping
require("gridExtra") #Arrange Plots for output
require("plyr") #for ddply
require("dplyr")
require("utils")
library("tidyverse")
```

Temperature treatments during experiment.  

# Larval temperature treatments  

Load temperature files for ambient and cold sump during larval rearing and calibrate to calibration file.  
```{r}
#Load calibration data, extract intercept and slope to calibrate individual loggers
calib.data.T <- read.csv("Temperature/Temp_Logger_Calibration.csv", header=TRUE) #load data with a header, separated by commas, with NA as NA
calib.T <- as.data.frame(t(apply(calib.data.T[2:21], 2, function(y)(coef(lm(y ~ calib.data.T$SN_10779069)))))) #apply linear model to obtain calibration equations
colnames(calib.T) <- c("Intercept",  "Slope") #rename columns

#SUMP only
#Load data
AmbSump<-read.csv("Temperature/Ambient_Sump_10779069.csv",header=TRUE, sep=",", na.strings="NA") #this logger is the standard, no change
#Intercept: -2.757311e-14     slope= 1.0000000, line equation y=1  x +-2.757311e-14
AmbSump$calibratedT<-(AmbSump$Temp*1)+-2.757311e-14
names(AmbSump)[3] <- "AmbientSump" #rename calibrated col to tank name
AmbSump<- AmbSump[ -c(2) ] #drop uncalibrated col

ColdSump<-read.csv("Temperature/cold_sump_10487935.csv",header=TRUE, sep=",", na.strings="NA")
#Intercept: -5.645523e-02   Slope: 1.0004740, line equation y=1.0004740x + -5.645523e-02
ColdSump$calibratedT<-(ColdSump$Temp*1.0004740)+-5.645523e-02
names(ColdSump)[3] <- "ColdSump" #rename calibrated col to tank name
ColdSump<- ColdSump[ -c(2) ] #drop uncalibrated col
```


Plot ambient and cold sumps for reference. Note that temperature data presented was obtained from individual temperature loggers in conical vessels. 

```{r}
#combine Amb and cold sump for plotting
SUMPdf<-merge(AmbSump,ColdSump, by="Date.Time") #merge and keep only time points over lap
SUMPdf<-melt(SUMPdf,id=c("Date.Time"))
SUMPdf$Date.Time<-as.character(SUMPdf$Date.Time)
SUMPdf$Date <- format(as.POSIXct(strptime(SUMPdf$Date.Time,"%m/%d/%Y %H:%M",tz="")) ,format = "%m/%d") #to get daily means set up a date only col

SUMP_mean <- ddply(SUMPdf, c("variable", "Date"), summarise, 
                          N    = length(value[!is.na(value)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                          mean = mean(value, na.rm=TRUE), #calculate mean of response variable, removing NA's
                          sd   = sd(value, na.rm=TRUE), #calculate standard deviation
                          se   = sd / sqrt(N), #calculate standard error
                          max = max(value, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
SUMP_mean

SUMP_plot<-ggplot(data=SUMP_mean, aes(x=Date, y=mean, group = variable, color=variable)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),
                width=0.1)+
  geom_point(aes(shape=variable), size=3)+
  geom_line(aes(color=variable, linetype=variable))+
  scale_color_manual(values=c("#000000","#0072B2"))+
  #scale_linetype_manual(values=c("solid", "dotted"))+
 # scale_color_manual(values=c("black", "black","black"))+
  ylab("Temp") +
  xlab("Time") + #Label the X Axis
  #ylim(0, 100) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("SUMP TEMP")+
  theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));SUMP_plot
```

Cold and ambient temperature sumps were distinct temperature regimes.  

Plot temperature data from loggers in conicals and view averages of each treatment (Script and data from Hancock et al. 2021).  

```{r}
library(readxl);library(tidyverse);library(lubridate);library(scales);library(plotrix)
hot<-read_excel("Temperature/conicals.xlsx",sheet="Hot");str(data)
ambient<-read_excel("Temperature/conicals.xlsx",sheet="Ambient");str(ambient)
cold<-read_excel("Temperature/conicals.xlsx",sheet="Cold");str(cold)

data<-full_join(full_join(hot,ambient,by="Date"),cold,by="Date")%>%
     dplyr::rename(Hot=Temp.x,Ambient=Temp.y)%>%
     gather(Treatment,Temp,-Date)%>%
     filter(Date>as.POSIXct("2019-07-06")&Date<as.POSIXct("2019-07-9"))
data$Treatment <- factor(data$Treatment,levels = c("Cold", "Ambient", "Hot"))

conical_plot<-ggplot(data)+geom_point(aes(Date,Temp,color=Treatment),alpha=0.3,size=1)+
     theme_classic()+
     scale_color_manual(values=c("darkblue","lightblue","red"), labels=c("Cool", "Ambient", "Warm"))+
     geom_smooth(aes(Date,Temp,color=Treatment),se=FALSE,span=0.15,size=0.5)+
     scale_x_datetime(date_breaks="1 day",minor_breaks=waiver(),labels=date_format("%m-%d"))+
     #scale_y_continuous(breaks=seq(23,32,1))+
     ylim(22.5, 32.5)+
     ylab("Temperature (°C)")+
     theme(legend.key.size = unit(1,"line"))+
    theme(legend.position="none")+
    theme(legend.title=element_text(size=10))+
    theme(plot.margin = margin(0.5, 0.1, 0.1, 0.1, "cm"))+
    theme(legend.text=element_text(size=9)); conical_plot

#ggsave("Figures/concial_plot.png", plot=conical_plot, w=5, h=3, units="in", dpi=300)

data%>%group_by(Treatment)%>%summarise(mean=mean(Temp),sd=sd(Temp),se=std.error(Temp))


#create a legend for later 
#build a plot for the legend
legend_plot<-conical_plot + 
  theme(legend.position="right")

# extract the legend from one of the plots
legend <- get_legend(
  # create some space to the left of the legend
  legend_plot)

#+ theme(legend.box.margin = margin(1,1,1,1))
#)

```


# Temperatures during larval settlement and recruit survival  

Display temperatures in tanks during larval settlement and recruit survival.  

```{r}
#TANKS ONLY
Tank1<-read.csv("Temperature/Tank1_5Sept_10779060.csv",header=TRUE, sep=",", na.strings="NA")
#Intercept: -5.899138e-02     Slope: 1.0015297 , line equation y=1.0015297 x + -5.899138e-02  
Tank1$calibratedT<-(Tank1$Temp*1.0015297)+-5.899138e-02
names(Tank1)[4] <- "Tank1" #rename calibrated col to tank name
Tank1<- Tank1[ -c(2) ] #drop uncalibrated col

Tank2<-read.csv("Temperature/Tank2_5Sept_10779068.csv",header=TRUE, sep=",", na.strings="NA")
#Intercept: 6.285066e-02  Slope: 0.9975927   , line equation y=0.9975927   x + 6.285066e-02  
Tank2$calibratedT<-(Tank2$Temp*1.0015297)+6.285066e-02 
names(Tank2)[4] <- "Tank2" #rename calibrated col to tank name
Tank2<- Tank2[ -c(2) ] #drop uncalibrated col

Tank3<-read.csv("Temperature/Tank3_5Sept_10487935.csv",header=TRUE, sep=",", na.strings="NA")
#Intercept: -5.645523e-02  Slope: 1.0004740   , line equation y=1.0004740   x +-5.645523e-02  
Tank3$calibratedT<-(Tank3$Temp*1.0004740)+-5.645523e-02
names(Tank3)[4] <- "Tank3" #rename calibrated col to tank name
Tank3<- Tank3[ -c(2) ] #drop uncalibrated col

Tank4<-read.csv("Temperature/Tank4_5Sept_10779069.csv",header=TRUE, sep=",", na.strings="NA")
#Intercept: -2.757311e-14     slope= 1.0000000, line equation y=1  x +-2.757311e-14
Tank4$calibratedT<-(Tank4$Temp*1)+-2.757311e-14
names(Tank4)[4] <- "Tank4" #rename calibrated col to tank name
Tank4<- Tank4[ -c(2) ] #drop uncalibrated col

#Join dataframes together  
Tankdf<-right_join(Tank1, Tank2, by=c("DateTime", "Phase"))
Tankdf<-right_join(Tankdf, Tank3, by=c("DateTime", "Phase"))
Tankdf<-right_join(Tankdf, Tank4, by=c("DateTime", "Phase"))

Tankdf$DateTime<-as.character(as.POSIXct(Tankdf$DateTime, format="%m/%d/%y %H:%M"))

#Tankdf$Date.Time<-as.character(Tankdf$DateTime)
#Tankdf$Date <- format(as.POSIXct(strptime(Tankdf$DateTime,"%m/%d/%Y %H:%M",tz="")) ,format = "%m/%d") #to get daily means set up a date only col
#Tankdf$Hour <- format(as.POSIXct(strptime(Tankdf$DateTime,"%m/%d/%Y %H:%M",tz="")) ,format = "%H") #to get daily means set up a date only col
#Tankdf$DateHour <- format(as.POSIXct(strptime(Tankdf$DateTime,"%m/%d/%Y %H:%M",tz="")) ,format ="%m/%d %H") #to get daily means set up a date only col

```

Separate settlement and survival phases and assign temperature treatment names.  

```{r}
settlement_temps<-Tankdf%>%
  filter(Phase=="Settlement")%>%
  gather(Tank, Temp, Tank1:Tank4)

settlement_temps$Treatment<-"Ambient"

survival_temps<-Tankdf%>%
  filter(Phase=="Survival")%>%
  gather(Tank, Temp, Tank1:Tank4)%>%
  mutate(Treatment = case_when(Tank=="Tank1" ~ "High", 
                               Tank=="Tank2" ~ "Ambient", 
                               Tank=="Tank3" ~ "Ambient", 
                               Tank=="Tank4" ~ "High"))
```


## Larval settlement  

Plot temperatures during larval settlement and calculate mean values.  
```{r}
settlement_plot<-ggplot(settlement_temps)+geom_point(aes(as.POSIXct(DateTime),Temp,color=Treatment),alpha=0.3,size=1)+
     theme_classic()+
     scale_color_manual(values=c("lightblue"))+
     geom_smooth(aes(as.POSIXct(DateTime),Temp,color=Treatment),se=FALSE,span=0.15,size=0.5)+
     scale_x_datetime(date_breaks="4 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
     #scale_y_continuous(breaks=seq(23,32,1))+
     ylim(22.5, 32.5)+
     xlab("Date")+
     ylab("Temperature (°C)")+
     theme(legend.key.size = unit(0.5,"line"))+
    theme(plot.margin = margin(0.5, 0.1, 0.1, 0.1, "cm"))+
    theme(legend.position="none"); settlement_plot

#ggsave("Figures/settlement_temps.png", plot=settlement_plot, w=5, h=3, units="in", dpi=300)

settlement_temps%>%group_by(Treatment)%>%summarise(mean=mean(Temp), sd=sd(Temp), se=std.error(Temp))

```


## Recruit survival

Plot temperatures during recruit survival and calculate mean values.  
```{r}
recruit_plot<-ggplot(survival_temps)+geom_point(aes(as.POSIXct(DateTime),Temp,color=Treatment),alpha=0.3,size=1)+
     theme_classic()+
     scale_color_manual(values=c("lightblue", "red"))+
     geom_smooth(aes(as.POSIXct(DateTime),Temp,color=Treatment),se=FALSE,span=0.15,size=0.5)+
     scale_x_datetime(date_breaks="8 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
     #scale_y_continuous(breaks=seq(23,32,1))+
     ylim(22.5, 32.5)+
     xlab("Date")+
     ylab("Temperature (°C)")+
     theme(legend.key.size = unit(0.5,"line"))+
    theme(plot.margin = margin(0.5, 0.1, 0.1, 0.1, "cm"))+
    theme(legend.position="none"); recruit_plot

#ggsave("Figures/recruit_temps.png", plot=recruit_plot, w=5, h=3, units="in", dpi=300)

survival_temps%>%group_by(Treatment)%>%summarise(mean=mean(Temp), sd=sd(Temp), se=std.error(Temp))

```


# Generate temperature figure  

```{r}
library(cowplot)

temp_panel<-plot_grid(conical_plot, settlement_plot, recruit_plot, labels = c("a. Larval Exposure", "b. Larval Settlement", "c. Recruit Survival"), ncol=3, nrow=1, rel_heights= c(1,1,1), rel_widths = c(1,1,1), label_size = 10, label_y=1, align="vh")

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
temp_plot_legend<-plot_grid(temp_panel, legend, rel_widths = c(6, .6), ncol=2, nrow=1)
 
ggsave(filename="Figures/Full_Temp_Figure.png", plot=temp_plot_legend, dpi=500, width=12, height=2, units="in")
```












